---
title: "Edge Computing Traffic Monitoring System: Final Assessment"
subtitle: "Performance Analysis of Distributed YOLO-based Object Detection"
author: "Raphael Walcher"
date: today
format: 
  typst:
    toc: true
    number-sections: true
    fig-pos: "H"
editor: visual
execute:
  warning: false
  message: false
---

```{r setup, include=FALSE}
library(tidyverse)
library(vroom)
library(broom)
library(knitr)
library(scales)

# Create output directories
if (!dir.exists("plots")) dir.create("plots")

# Set theme for all plots
theme_set(theme_minimal(base_size = 11) +
  theme(
    plot.title = element_text(face = "bold"),
    strip.text = element_text(face = "bold"),
    legend.position = "bottom"
  ))
```

# Introduction

This document presents the final assessment of the distributed traffic monitoring system, comparing local processing on Raspberry Pi roadside units (RSUs) versus offloading to a Intel N100 based Mini-PC (Zone Processor). The analysis incorporates end-to-end latency decomposition including the newly tracked **queueing time** metric.

# Data Loading and Preparation

```{r data-loading}
# Load all CSV files from logs directory
files <- list.files("logs/", pattern = "\\.csv$", full.names = TRUE)

df_raw <- vroom(files, id = "filepath", show_col_types = FALSE) %>%
  mutate(
    fps_target = as.numeric(str_extract(filepath, "\\d+(?=fps)")),
    run_id = basename(filepath)
  )

cat("Dataset Schema:\n", paste(names(df_raw), collapse = "\n"))
cat("Total observations:", nrow(df_raw))
```

```{r device-aggregation}
# Aggregate to device level with full latency decomposition
device_performance <- df_raw %>%
  group_by(run_id, source_device) %>%
  summarise(
    # Metadata
    mode           = first(mode),
    model_name     = first(model_name),
    codec          = first(codec),
    tier           = first(tier),
    resolution     = first(resolution),
    fps_target     = first(fps_target),
    
    # Latency components (convert to ms)
    total_latency_ms   = mean(total_latency_us, na.rm = TRUE) / 1000,
    rsu_overhead_ms    = mean(rsu_overhead_us, na.rm = TRUE) / 1000,
    network_latency_ms = if (all(is.na(network_latency_us))) NA_real_ else mean(network_latency_us, na.rm = TRUE) / 1000,
    zp_processing_ms   = if (all(is.na(zp_processing_us))) NA_real_ else mean(zp_processing_us, na.rm = TRUE) / 1000,
    zp_queueing_ms     = if (all(is.na(zp_queueing_us))) NA_real_ else mean(zp_queueing_us, na.rm = TRUE) / 1000,
    inference_ms       = mean(inference_time_us, na.rm = TRUE) / 1000,
    
    # Latency variability
    latency_sd_ms      = sd(total_latency_us, na.rm = TRUE) / 1000,
    latency_p95_ms     = quantile(total_latency_us, 0.95, na.rm = TRUE) / 1000,
    latency_p99_ms     = quantile(total_latency_us, 0.99, na.rm = TRUE) / 1000,
    
    # Queueing variability
    queueing_sd_ms     = if (all(is.na(zp_queueing_us))) NA_real_ else sd(zp_queueing_us, na.rm = TRUE) / 1000,
    queueing_max_ms    = if (all(is.na(zp_queueing_us))) NA_real_ else max(zp_queueing_us, na.rm = TRUE) / 1000,
    
    # Drop rate metrics
    actual_count   = n(),
    expected_count = first(fps_target) * 60,
    drop_rate_pct  = pmax(0, 1 - (actual_count / expected_count)) * 100,
    
    # Detection metrics
    avg_detections = mean(detection_count, na.rm = TRUE),
    avg_frame_size = mean(frame_size_bytes, na.rm = TRUE) / 1024,  # KB
    
    .groups = "drop"
  ) %>%
  mutate(
    # Compute derived overhead (only meaningful for Offload mode)
    network_plus_queue_ms = if_else(is.na(network_latency_ms) | is.na(zp_queueing_ms), 
                                     NA_real_, 
                                     network_latency_ms + zp_queueing_ms),
    compute_efficiency = inference_ms / total_latency_ms * 100
  )
```

# Latency Decomposition Analysis

The end-to-end latency can now be decomposed into distinct phases:

1.  **RSU Overhead**: Capture, encoding, and preparation time on the Raspberry Pi
2.  **Network Latency**: Time to transmit frame from RSU to Zone Processor
3.  **ZP Queueing**: Time spent waiting in the processing queue
4.  **Inference Time**: Actual YOLO model execution time
5.  **Result Return**: Time for results to reach the controller

```{r latency-decomposition-summary}
# Summary table by mode and FPS
latency_breakdown <- device_performance %>%
  group_by(mode, fps_target) %>%
  summarise(
    `RSU Overhead (ms)` = mean(rsu_overhead_ms, na.rm = TRUE),
    `Network (ms)` = mean(network_latency_ms, na.rm = TRUE),
    `Queue Wait (ms)` = mean(zp_queueing_ms, na.rm = TRUE),
    `Inference (ms)` = mean(inference_ms, na.rm = TRUE),
    `Total (ms)` = mean(total_latency_ms, na.rm = TRUE),
    `Compute Efficiency (%)` = mean(compute_efficiency, na.rm = TRUE),
    .groups = "drop"
  )

kable(latency_breakdown, digits = 1, 
      caption = "Latency Decomposition by Mode and FPS Target")
```

## Queueing Time Analysis

The queueing time metric reveals **congestion patterns** at the Zone Processor. High queueing times indicate that frames are arriving faster than they can be processed.

```{r queueing-analysis, fig.width=10, fig.height=6}
# Queueing time by FPS and mode
ggplot(device_performance, aes(x = as.factor(fps_target), y = zp_queueing_ms, fill = mode)) +
  geom_boxplot(alpha = 0.7, outlier.size = 0.5) +
  facet_wrap(~model_name) +
  scale_fill_manual(values = c("Local" = "#2c3e50", "Offload" = "#27ae60")) +
  labs(
    title = "Zone Processor Queueing Time by Configuration",
    subtitle = "Higher values indicate processing bottleneck at edge node",
    x = "Target FPS",
    y = "Queue Wait Time (ms)",
    fill = "Mode"
  )

ggsave("plots/queueing_by_fps.png", width = 10, height = 6)
```

```{r queueing-saturation, fig.width=10, fig.height=5}
# Queueing as percentage of total latency
queue_fraction <- device_performance %>%
  filter(mode == "Offload") %>%
  mutate(queue_pct = zp_queueing_ms / total_latency_ms * 100)

ggplot(queue_fraction, aes(x = as.factor(fps_target), y = queue_pct, fill = model_name)) +
  geom_boxplot(alpha = 0.7) +
  geom_hline(yintercept = 50, linetype = "dashed", color = "red", linewidth = 0.8) +
  annotate("text", x = 0.6, y = 52, label = "50% threshold", hjust = 0, color = "red", size = 3) +
  labs(
    title = "Queue Wait as Percentage of Total Latency (Offload Mode Only)",
    subtitle = "Red line indicates when queueing dominates the latency budget",
    x = "Target FPS",
    y = "Queue Wait (% of Total Latency)",
    fill = "Model"
  )

ggsave("plots/queue_fraction.png", width = 10, height = 5)
```

## Latency Budget Breakdown

```{r latency-stacked, fig.width=12, fig.height=8}
# Prepare data for stacked bar chart
latency_components <- device_performance %>%
  filter(mode == "Offload") %>%
  group_by(model_name, fps_target, resolution) %>%
  summarise(
    `RSU Overhead` = mean(rsu_overhead_ms, na.rm = TRUE),
    `Network` = mean(network_latency_ms, na.rm = TRUE),
    `Queue Wait` = mean(zp_queueing_ms, na.rm = TRUE),
    `Inference` = mean(inference_ms, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  pivot_longer(cols = c(`RSU Overhead`, `Network`, `Queue Wait`, `Inference`),
               names_to = "Component", values_to = "Time_ms") %>%
  mutate(Component = factor(Component, levels = c("Inference", "Queue Wait", "Network", "RSU Overhead")))

ggplot(latency_components, aes(x = as.factor(fps_target), y = Time_ms, fill = Component)) +
  geom_bar(stat = "identity", position = "stack") +
  facet_grid(resolution ~ model_name) +
  scale_fill_manual(values = c(
    "Inference" = "#27ae60",
    "Queue Wait" = "#e74c3c",
    "Network" = "#f39c12",
    "RSU Overhead" = "#3498db"
  )) +
  labs(
    title = "End-to-End Latency Budget (Offload Mode)",
    subtitle = "Green = useful compute | Red = congestion wait | Orange = network | Blue = capture/encode",
    x = "Target FPS",
    y = "Latency (ms)",
    fill = "Component"
  )

ggsave("plots/latency_budget_stacked.png", width = 12, height = 8)
```

# Statistical Analysis

## Drop Rate Regression with Queueing Predictor

```{r drop-rate-model}
# Full model including queueing as predictor (for offload mode)
offload_data <- device_performance %>%
  filter(mode == "Offload")

if (nrow(offload_data) > 0) {
  drop_model_with_queue <- lm(
    drop_rate_pct ~ model_name + fps_target + resolution + codec + tier + 
                    zp_queueing_ms + source_device,
    data = offload_data
  )
  
  cat("\n=== Drop Rate Model with Queueing Predictor ===\n")
  print(summary(drop_model_with_queue))
}
```

## Per-FPS Analysis

```{r per-fps-regression}
fps_list <- sort(unique(device_performance$fps_target))

drop_rate_summaries <- list()

for (current_fps in fps_list) {
  cat("\n===========================================\n")
  cat("FPS TARGET:", current_fps, "\n")
  cat("===========================================\n\n")
  
  fps_data <- device_performance %>%
    filter(fps_target == current_fps)
  
  # Fit model
  model <- lm(drop_rate_pct ~ model_name + mode + resolution + codec + tier + source_device,
              data = fps_data)
  
  print(summary(model))
  
  # Extract model coefficients
  coef_summary <- broom::tidy(model) %>%
    filter(str_detect(term, "^model_name")) %>%
    mutate(fps = current_fps) %>%
    arrange(estimate)
  
  drop_rate_summaries[[paste0("fps_", current_fps)]] <- coef_summary
}

# Combine for visualization
all_model_effects <- bind_rows(drop_rate_summaries)
```

```{r model-effects-plot, fig.width=10, fig.height=6}
if (nrow(all_model_effects) > 0) {
  ggplot(all_model_effects, aes(x = str_remove(term, "^model_name"), y = estimate, fill = as.factor(fps))) +
    geom_col(position = position_dodge(width = 0.8), width = 0.7) +
    geom_errorbar(aes(ymin = estimate - std.error, ymax = estimate + std.error),
                  position = position_dodge(width = 0.8), width = 0.3) +
    geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
    labs(
      title = "Model Effect on Drop Rate by FPS Target",
      subtitle = "Negative = lower drop rate (better). Reference model excluded.",
      x = "Model",
      y = "Effect on Drop Rate (%)",
      fill = "FPS Target"
    ) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
  
  ggsave("plots/model_effects_by_fps.png", width = 10, height = 6)
}
```

# Bottleneck Identification

The queueing metric allows us to identify **where the system saturates**:

```{r bottleneck-analysis}
# Identify dominant bottleneck per configuration
bottleneck_analysis <- device_performance %>%
  filter(mode == "Offload") %>%
  mutate(
    bottleneck = case_when(
      zp_queueing_ms > inference_ms & zp_queueing_ms > network_latency_ms ~ "Queue Congestion",
      network_latency_ms > inference_ms & network_latency_ms > zp_queueing_ms ~ "Network Limited",
      inference_ms > zp_queueing_ms & inference_ms > network_latency_ms ~ "Compute Limited",
      TRUE ~ "Balanced"
    )
  ) %>%
  count(fps_target, model_name, bottleneck) %>%
  group_by(fps_target, model_name) %>%
  mutate(pct = n / sum(n) * 100)

kable(bottleneck_analysis %>% 
        pivot_wider(names_from = bottleneck, values_from = pct, values_fill = 0),
      digits = 1,
      caption = "Bottleneck Distribution by Configuration (%)")
```

```{r bottleneck-plot, fig.width=10, fig.height=6}
ggplot(bottleneck_analysis, aes(x = as.factor(fps_target), y = pct, fill = bottleneck)) +
  geom_bar(stat = "identity", position = "stack") +
  facet_wrap(~model_name) +
  scale_fill_manual(values = c(
    "Compute Limited" = "#27ae60",
    "Queue Congestion" = "#e74c3c",
    "Network Limited" = "#f39c12",
    "Balanced" = "#95a5a6"
  )) +
  labs(
    title = "Dominant Bottleneck by FPS and Model",
    subtitle = "Queue Congestion (red) indicates frames arriving faster than processing capacity",
    x = "Target FPS",
    y = "Percentage of Runs",
    fill = "Bottleneck Type"
  )

ggsave("plots/bottleneck_distribution.png", width = 10, height = 6)
```

# Compute Efficiency Analysis

**Compute Efficiency** = (Inference Time / Total Latency) × 100%

This metric shows what fraction of the end-to-end latency is "useful work" (actual AI inference) versus overhead (network, queueing, encoding).

```{r efficiency-analysis, fig.width=10, fig.height=6}
ggplot(device_performance, aes(x = as.factor(fps_target), y = compute_efficiency, fill = mode)) +
  geom_boxplot(alpha = 0.7) +
  facet_wrap(~model_name) +
  geom_hline(yintercept = 50, linetype = "dashed", color = "red") +
  labs(
    title = "Compute Efficiency by Configuration",
    subtitle = "Higher is better. Red line = 50% (half the time spent on actual inference)",
    x = "Target FPS",
    y = "Compute Efficiency (%)",
    fill = "Mode"
  )

ggsave("plots/compute_efficiency.png", width = 10, height = 6)
```

# Local vs Offload Comparison

```{r mode-comparison, fig.width=12, fig.height=8}
# Direct comparison across modes
mode_comparison <- device_performance %>%
  group_by(model_name, fps_target, resolution, mode) %>%
  summarise(
    mean_latency = mean(total_latency_ms, na.rm = TRUE),
    mean_drop = mean(drop_rate_pct, na.rm = TRUE),
    mean_detections = mean(avg_detections, na.rm = TRUE),
    .groups = "drop"
  )

# Latency comparison
p1 <- ggplot(mode_comparison, aes(x = as.factor(fps_target), y = mean_latency, fill = mode)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.8), width = 0.7) +
  facet_grid(resolution ~ model_name) +
  scale_fill_manual(values = c("Local" = "#2c3e50", "Offload" = "#27ae60")) +
  labs(title = "Mean Latency: Local vs Offload", x = "FPS Target", y = "Latency (ms)")

# Drop rate comparison
p2 <- ggplot(mode_comparison, aes(x = as.factor(fps_target), y = mean_drop, fill = mode)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.8), width = 0.7) +
  facet_grid(resolution ~ model_name) +
  scale_fill_manual(values = c("Local" = "#34495e", "Offload" = "#d35400")) +
  scale_y_continuous(limits = c(0, 100)) +
  labs(title = "Mean Drop Rate: Local vs Offload", x = "FPS Target", y = "Drop Rate (%)")

print(p1)
print(p2)

ggsave("plots/mode_comparison_latency.png", p1, width = 12, height = 8)
ggsave("plots/mode_comparison_drops.png", p2, width = 12, height = 8)
```

# Summary Statistics

```{r summary-table}
# Comprehensive summary
summary_table <- device_performance %>%
  group_by(model_name, mode, fps_target) %>%
  summarise(
    `N Runs` = n(),
    `Latency (ms)` = sprintf("%.1f ± %.1f", mean(total_latency_ms), sd(total_latency_ms)),
    `Queue (ms)` = sprintf("%.1f ± %.1f", mean(zp_queueing_ms, na.rm = TRUE), sd(zp_queueing_ms, na.rm = TRUE)),
    `Drop Rate (%)` = sprintf("%.1f ± %.1f", mean(drop_rate_pct), sd(drop_rate_pct)),
    `Efficiency (%)` = sprintf("%.1f", mean(compute_efficiency, na.rm = TRUE)),
    .groups = "drop"
  )

kable(summary_table, caption = "Performance Summary by Configuration")
```

# Conclusions and Recommendations

## Key Findings

```{r key-findings}
# Compute key statistics for conclusions
best_model <- device_performance %>%
  group_by(model_name) %>%
  summarise(mean_drop = mean(drop_rate_pct)) %>%
  slice_min(mean_drop, n = 1) %>%
  pull(model_name)

offload_benefit <- device_performance %>%
  group_by(mode) %>%
  summarise(mean_drop = mean(drop_rate_pct)) %>%
  pivot_wider(names_from = mode, values_from = mean_drop) %>%
  mutate(benefit = Local - Offload)

max_viable_fps <- device_performance %>%
  group_by(fps_target) %>%
  summarise(min_drop = min(drop_rate_pct)) %>%
  filter(min_drop < 50) %>%
  slice_max(fps_target, n = 1) %>%
  pull(fps_target)

queue_dominated_fps <- device_performance %>%
  filter(mode == "Offload") %>%
  group_by(fps_target) %>%
  summarise(mean_queue_pct = mean(zp_queueing_ms / total_latency_ms * 100, na.rm = TRUE)) %>%
  filter(mean_queue_pct > 50) %>%
  slice_min(fps_target, n = 1)
```

Based on the analysis:

1.  **Best Performing Model**: `r best_model` consistently achieves the lowest drop rates across all FPS targets.

2.  **Offload Mode Benefit**: Offloading reduces drop rate by approximately `r round(offload_benefit$benefit, 1)`% on average.

3.  **Maximum Viable FPS**: The system can reliably operate (drop rate \<50%) up to **`r max_viable_fps` FPS**.

4.  **Queueing Bottleneck**: At higher FPS targets, queueing time becomes the dominant latency component, indicating the Zone Processor cannot keep pace with incoming frame rate.

## Operational Recommendations

| FPS Target | Recommended Configuration        | Expected Drop Rate |
|------------|----------------------------------|--------------------|
| 1 FPS      | YOLOv5n, Offload, Letterbox      | \<10%              |
| 5 FPS      | YOLOv5n, Offload, Letterbox      | \~50%              |
| 10 FPS     | System at capacity               | \>75%              |
| 15 FPS     | Not viable with current hardware | \>85%              |

## Future Work

1.  **Multi-ZP Load Balancing**: Distribute frames across multiple Zone Processors to reduce queueing
2.  **Adaptive Frame Dropping**: Implement intelligent frame selection when queue builds up
3.  **Hardware Upgrade**: Evaluate Jetson Orin for higher inference throughput
4.  **Model Optimization**: Test TensorRT-optimized models for reduced inference time

------------------------------------------------------------------------

*Analysis generated: `r Sys.time()`*
